---
title: "p8105_hw2_jg4891"
output: github_document
date: "2025-09-24"
---

```{r setup}
library(tidyverse)
library(ggridges)
options(tibble.print_min = 5)

```
## Problem 1
```{r pol_month}
pols_df <- read_csv("fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = '-') %>%
  mutate(
    year  = as.character(year),
    month = as.integer(month),       # "01" -> 1, "02" -> 2
    month = month.name[month],       # full month names
    month = as.character(month),
    president = case_when(
      prez_gop == 1 ~ "Republican",
      prez_dem == 1 ~ "Democratic"
    )
  ) %>%
  select(-prez_gop, -prez_dem, -day)
```
```{r snp}
snp_df <- read_csv("fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = '/') %>%
  mutate(
    month = as.integer(month),
    year  = as.integer(year),
    year  = ifelse(year < 50, 2000 + year, 1900 + year),
    month = month.name[month],     # full month names
    year  = as.character(year),
    month = as.character(month)
  ) %>%
  select(-day) %>%
  select(year, month, everything())
```
```{r unemploy}
unemploy_df <- read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
janitor::clean_names() %>% 
pivot_longer(
    cols = jan:dec,                  # columns to pivot
    names_to = "month",              # new column with month names
    values_to = "unemployment_percent" # new column with values
  ) %>% 
  
  mutate(
    month = month.name[match(tolower(month), tolower(month.abb))],
    month = factor(month, levels = month.name, ordered = TRUE),
    year  = as.character(year),
    month = as.character(month)
  )
```
```{r merge}

ft_df = left_join(pols_df, snp_df, by = c("year", "month"))
Sd_df = left_join(ft_df, unemploy_df, by = c("year", "month"))
print(Sd_df)

```

Pols_df dataset contains the number of national politicians who are democratic or republican at any given time from 1947 Jan to 2015 June.
Snp_df dataset includes closing values of the S&P stock index from 1950 Jan to 2015 July.
Unemploy_df dataset includes unemployment percentages from  1948 Jan to 2015 Dec. 
This merged dataset reflects number of republican governors and senators, democratic governors and senators, and president's party, and unemployment percentage from 1947 to 2015.

## Problem 2

```{r}
  mr_df = 
  readxl::read_excel("Trash.xlsx", sheet="Mr. Trash Wheel", range = "A2:N653",na = c("NA", "", "."))|>
  filter(!is.na(Dumpster)) |> 
  janitor::clean_names() |> 
  mutate(sheet = "Mr. Trash") |> 
  mutate(sports_balls = as.integer(round(sports_balls)))

```


```{r}
  pr_df = 
  readxl::read_excel("Trash.xlsx", sheet="Professor Trash Wheel", range = "A2:M120",na = c("NA", "", "."))|>
  filter(!is.na(Dumpster)) |> 
  janitor::clean_names() |> 
  mutate(sheet = "Professor Trash")

```

```{r}
  gw_df = 
  readxl::read_excel("Trash.xlsx", sheet="Gwynnda Trash Wheel", range = "A2:L265",na = c("NA", "", "."))|>
   filter(!is.na(Dumpster)) |> 
  janitor::clean_names() |> 
  mutate(sheet = "Gwynnda Trash")

```


```{r}

df_tidy = bind_rows(gw_df %>% mutate(year = as.numeric(year)), mr_df%>% mutate(year = as.numeric(year)), pr_df%>% mutate(year = as.numeric(year))) |> 
  janitor::clean_names() |> 
  select(sheet, everything())


pf_wt <- df_tidy %>% 
  filter(sheet == "Professor Trash") %>% 
  summarize(total_weight = sum(weight_tons, na.rm = TRUE)) %>% 
  pull(total_weight)

gw_cb <- df_tidy %>% 
  filter(sheet == "Gwynnda Trash", year=="2022", month=="June") %>% 
  summarize(total_weight = sum(cigarette_butts, na.rm = TRUE)) %>% 
  pull(total_weight)
```


Variables definition: Dumpster shows the code for each dumpster in numeric value. Sheet is the dataset that each dumpster belong to, such as Mr. Trash, Professor Trash or Gynnda Trash. Weight_tons is the weight(tons) for each dumpster. Plastic bottles shows up the number of plastic bottles in each dumpster. Homes_powered is the energy that trash generate for home.

In total, we have `r nrow(df_tidy)` observations. 
In Professor Trash Wheel dataset, we have total weight of trash of  `r pf_wt` tons.
In Gwynnda dataset, in June 2022, we have a total number of cigarette butts collected is `r gw_cb`. 

## Problem 3
```{r warning=FALSE}
z1_df = read_csv("zillow_data/Zip_NYC.csv", na = c("NA", "", ".")) |> 
  janitor::clean_names() |> 
  rename(zip_code=region_name) |> 
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "time", 
    values_to = "Index") 

z2_df <- read_csv("zillow_data/Zipcodes.csv",na = c("NA", "", ".")) |> 
  janitor::clean_names() 

m2_df = left_join(z1_df, z2_df, by="zip_code") |> 
  filter(!is.na(Index)) |> 
  select(-region_type,-state_name,-county_name) |> 
  relocate(time, size_rank, Index,  state, state_fips,  city, zip_code, county, county_fips, county_code, region_id, metro, neighborhood, file_date)

```

```{r missing zipcodes}
missing_zips <- setdiff(z2_df$zip_code, z1_df$zip_code)
print(missing_zips)
```
Rental prices fluctuated dramatically during the COVID-19 pandemic. For all available ZIP codes, compare rental prices in January 2021 to prices in January 2020. Make a table that shows the 10 ZIP codes (along with the borough and neighborhood) with largest drop in price from January 2020 to 2021. Comment.

```{r}
m3_df = m2_df |> 
  janitor::clean_names() |> 
mutate(
    date = ymd(str_remove(time, "^x")),                  # "x2020_01_31" -> Date
    yr   = year(date),
    mo   = month(date)
  ) %>%
  filter(mo == 1, yr %in% c(2020, 2021)) %>%             # keep Jan 2020 & Jan 2021
  # aggregate to one price per ZIP/year (use mean or median)
  group_by(zip_code, city, county, yr) %>%
  summarise(index_sum = sum(index, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = yr, values_from = index_sum, names_prefix = "y") %>%
  # keep ZIPs that have both years
  filter(!is.na(y2020), !is.na(y2021)) %>% 
mutate(
    drop = y2020 - y2021
  ) %>%
  arrange(desc(drop)) %>%   
  slice_head(n = 10) %>%
  mutate(
    y2020     = round(y2020, 1),
    y2021     = round(y2021, 1),
    drop      = round(drop, 1)
  ) %>%
  select(zip_code, city, county, y2020, y2021, drop)

# show results
print(m3_df, n = 10)




```



There are `r nrow(m2_df)` observations, `r length(unique(m2_df$zip_code))` unique zip codes, and `r length(unique(m2_df$neighborhood))` neighborhoods. 171 zipcodes were missing in the zillow dataset because it is too narrow, such as 10464, 10474, 10475 and so on.It is part of military housing so it is not open to public. 

Compared rental price in Jan 2020 to Jan 2021, 10007 area has the biggest fall of 912.6 dollars. 







